---
title: "QRMon anomalies detection"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    theme: spacelab
runtime: shiny
---

```{r setup, include=FALSE}
library(flexdashboard)
library(magrittr)
library(ggplot2)
library(QRMon)
library(OutlierIdentifiers)
```


Sidebar {.sidebar}
=====================================

```{r}
selectInput( inputId = "dataSpec", label = "Data:", choices = c( "DistributionData", "FinancialData", "TemperatureData" ), selected = "DistributionData" )

checkboxInput( inputId = "dateAxisQ", label = "Date axis:", value = FALSE )

hr()

textInput( inputId = "probs", label = "Probabilities:", value = "0.25, 0.5, 0.75")

sliderInput( inputId = "df", label = "Degrees of freedom:", min = 1, max = 120, step = 1, value = 12 )

hr()

checkboxInput( inputId = "relativeErrorsQ", label = "Use relative errors:", value = FALSE )

numericInput( inputId = "outlierThreshold", label = "Outliers threshold:", min = 0, max = 100, step = 0.05, value = 1 )

selectInput( inputId = "outlierIdentifier", 
             label = "Outlier identifier:", 
             choices = c( "Hampel", "SPLUSQuartile", "QuartileIdentifier"), 
             selected = "Hampel" )

```

Data and fit
=====================================

```{r}
dataset <- reactive({
  list( "DistributionData" = dfDistributionData, "FinancialData" = dfFinancialData, "TemperatureData" = dfTemperatureData )[[input$dataSpec]]
})

outlierIdentifierParameters <- reactive({
  list( 
    "Hampel" = HampelIdentifierParameters, 
    "SPLUSQuartile" = SPLUSQuartileIdentifierParameters, 
    "QuartileIdentifier" = QuartileIdentifierParameters
  )[[input$outlierIdentifier]]
})  

pointOutlierIdentifier <- reactive({
  function(x) TopOutlierIdentifier(data = x, identifier = outlierIdentifierParameters()) 
})

varianceOutlierIdentifier <- reactive({
  function(x) OutlierIdentifier(data = x, identifier = outlierIdentifierParameters()) 
})

lsProbs <- reactive({
  res <- strsplit( input$probs, ",|\\s")[[1]]
  as.numeric( res[nchar(res) > 0 ] )
})
  
qrObj <- reactive({

  QRMonUnit( setNames( dataset(), c("Regressor", "Value") ) ) %>%
    QRMonQuantileRegression( df = input$df, probabilities = lsProbs() )
})
```

### Summary

```{r}
renderPrint({
  cat( paste0("Number of points: ", nrow(dataset()), "\n"));
  summary(qrObj() %>% QRMonTakeData)
})
```


### Plot data and regression quantiles

```{r}
plotQR <- reactive({
  qrObj() %>% QRMonPlot( echoQ = FALSE, datePlotQ = input$dateAxisQ, dateOrigin = "1900-01-01" ) %>% QRMonTakeValue
})

renderPlot( expr = {
  print(plotQR())
})
```

### Plot errors

```{r}
renderPlot( expr = {
  print(qrObj() %>% QRMonErrorsPlot(relativeErrorsQ = input$relativeErrorsQ, datePlotQ = input$dateAxisQ ))
})
```

Point anomalies
=====================================

### Outliers by regression quantiles

```{r}
renderPlot( expr = {
  
  p <- 
    qrObj() %>% 
    QRMonOutliers %>% 
    QRMonOutliersPlot( echoQ = FALSE, plotRegressionCurvesQ = TRUE, datePlotQ = input$dateAxisQ, dateOrigin = "1900-01-01") %>% 
    QRMonTakeValue
  
  print(p)
  
})
```

### Anomalies by residuals using threshold

```{r}
dfPointOutliers <- reactive({
  qrObj() %>% 
    QRMonFindAnomaliesByResiduals( threshold = input$outlierThreshold, outlierIdentifier = NULL, relativeErrorsQ = input$relativeErrorsQ ) %>% 
    QRMonTakeValue
})

renderPlot( expr = {
  
  QRMonUnit( qrObj() %>% QRMonTakeData ) %>% 
    QRMonPlot( echoQ = FALSE, datePlotQ = input$dateAxisQ, dateOrigin = "1900-01-01" ) %>% 
    QRMonTakeValue +
    geom_point( data = dfPointOutliers(), 
                aes( x = if(input$dateAxisQ) { as.POSIXct(Regressor, origin = "1900-01-01") } else { Regressor }, 
                     y = Value ), 
                color = "red" ) 

})
```

### Anomalies by residuals using outlier identifier

```{r}
dfPointOutliersByFunc <- reactive({
  qrObj() %>% 
    QRMonFindAnomaliesByResiduals( threshold = NULL, outlierIdentifier = pointOutlierIdentifier(), relativeErrorsQ = input$relativeErrorsQ ) %>% 
    QRMonTakeValue
})

renderPlot( expr = {
  
  print(
    QRMonUnit( qrObj() %>% QRMonTakeData ) %>% 
      QRMonPlot( echoQ = FALSE, datePlotQ = input$dateAxisQ, dateOrigin = "1900-01-01" ) %>% 
      QRMonTakeValue +
      geom_point( data = dfPointOutliersByFunc(), 
                  aes( x = if(input$dateAxisQ) { as.POSIXct(Regressor, origin = "1900-01-01") } else { Regressor }, 
                       y = Value ), 
                  color = "red" ) 
  )
  
})
```


Variance anomalies
=====================================

### Regression quantiles for variance evaluation

```{r}
renderPlot( expr = {
  
  print(
    QRMonUnit( qrObj() %>% QRMonTakeData ) %>% 
      QRMonQuantileRegression( df = input$df, probabilities = c( min(lsProbs()), max(lsProbs()) ) ) %>% 
      QRMonPlot( echoQ = FALSE, datePlotQ = input$dateAxisQ, dateOrigin = "1900-01-01" ) %>% 
      QRMonTakeValue 
  )
  
})
```

### Conditional variance

```{r}
renderPlot( expr = {
  
  lsRes <- qrObj() %>% QRMonPredict( ) %>% QRMonTakeValue
  dfRes <- dplyr::bind_rows( lsRes, .id = "Prob")
  dfDiffs <- data.frame( Regressor = lsRes[[2]]["Regressor"], Diff = abs(lsRes[[2]]["Value"] - lsRes[[1]]["Value"] ) )

  lsOutlierPos <- varianceOutlierIdentifier()( dfDiffs$Value )
  
  if( length(lsOutlierPos) == 0 || sum(lsOutlierPos) == 0 ) { 
    ggplot( cbind( Type = "VarianceDiff", dfDiffs) ) +
      geom_point( aes( x = if(input$dateAxisQ) { as.POSIXct(Regressor, origin = "1900-01-01") } else { Regressor }, 
                       y = Value,  
                       color = Type)
      ) +
      xlab("Regressor")
  } else {
    ggplot( rbind( cbind( Type = "VarianceDiff", dfDiffs), 
                   cbind( Type = "Outliers", dfDiffs[lsOutlierPos, ,drop=F]) ) ) + 
      geom_point( aes( x = if(input$dateAxisQ) { as.POSIXct(Regressor, origin = "1900-01-01") } else { Regressor }, 
                       y = Value,  
                       color = Type)
      ) +
      xlab("Regressor")
  }
})
```

### Variance outliers

```{r}
dfVarOutliers <- reactive({
  qrObj() %>% 
    QRMonFindVarianceAnomalies( outlierIdentifier = varianceOutlierIdentifier(), positionsQ = FALSE ) %>% 
    QRMonTakeValue
})

renderPlot( expr = {
  
  print(
    QRMonUnit( qrObj() %>% QRMonTakeData ) %>% 
      QRMonPlot( echoQ = FALSE, datePlotQ = input$dateAxisQ, dateOrigin = "1900-01-01" ) %>% 
      QRMonTakeValue +
      geom_point( data = dfVarOutliers(), 
                  aes( x = if(input$dateAxisQ) { as.POSIXct(Regressor, origin = "1900-01-01") } else { Regressor }, 
                       y = Value ), 
                  color = "red" ) 
  )
})
```

References
====================================

#### In brief

The quantile regression workflows are made with the R package [`QRMon-R`](https://github.com/antononcube/QRMon-R),
which is based on [`quantreg`](https://cran.r-project.org/web/packages/quantreg/index.html).


#### Articles, books

[RK1] Roger Koenker, 
[Quantile Regression](https://books.google.com/books/about/Quantile_Regression.html?id=hdkt7V4NXsgC), 
Cambridge University Press, 2005.

[RK2] Roger Koenker,
["Quantile Regression in R: a vignette"](https://cran.r-project.org/web/packages/quantreg/vignettes/rq.pdf),
(2006),
[CRAN](https://cran.r-project.org/).

[AA1] Anton Antonov,
["A monad for Quantile Regression workflows"](https://github.com/antononcube/MathematicaForPrediction/blob/master/MarkdownDocuments/A-monad-for-Quantile-Regression-workflows.md),
(2018),
[MathematicaForPrediction at GitHub](https://github.com/antononcube/MathematicaForPrediction).

#### Packages

[RKp1] Roger Koenker,
[`quantreg`](https://cran.r-project.org/web/packages/quantreg/index.html),
[CRAN](https://cran.r-project.org/).

[AAp1] Anton Antonov,
[Quantile Regression Mathematica package](https://github.com/antononcube/MathematicaForPrediction/blob/master/QuantileRegression.m),
(2014),
[MathematicaForPrediction at GitHub](https://github.com/antononcube/MathematicaForPrediction).

[AAp2] Anton Antonov,
[Monadic Quantile Regression Mathematica package](https://github.com/antononcube/MathematicaForPrediction/blob/master/MonadicProgramming/MonadicQuantileRegression.m),
(2018),
[MathematicaForPrediction at GitHub](https://github.com/antononcube/MathematicaForPrediction).

[AAp3] Anton Antonov,
[`QuantileRegression`](https://resources.wolframcloud.com/FunctionRepository/resources/QuantileRegression),
(2019),
[Wolfram Function Repository](https://resources.wolframcloud.com/FunctionRepository/resources/QuantileRegression).

#### Repositories

[AAr1] Anton Antonov,
[QRMon-R](https://github.com/antononcube/QRMon-R),
(2019),
[GitHub/antononcube](https://github.com/antononcube).

[AAr2] Anton Antonov,
[DSL::English::QuantileRegressionWorkflows in Raku](https://github.com/antononcube/Raku-DSL-English-QuantileRegressionWorkflows),
(2020),
[GitHub/antononcube](https://github.com/antononcube).

#### Videos

[AAv1] Anton Antonov,
["Anomalies, Breaks, and Outliers Detection in Time Series"](https://www.youtube.com/watch?v=h_fLb6YU87c),
Wolfram Research Technology 2019 talk,
[Wolfram Research Inc. channel at YouTube](https://www.youtube.com/channel/UCJekgf6k62CQHdENWf2NgAQ).

[AAv2] Anton Antonov,
["Boston useR! QuantileRegression Workflows 2019-04-18"](https://www.youtube.com/watch?v=a_Dk25xarvE),
(2019),
[Anton Antonov at YouTube](https://www.youtube.com/channel/UC5qMPIsJeztfARXWdIw3Xzw).

[AAv3] Anton Antonov,
["useR! 2020: How to simplify Machine Learning workflows specifications"](https://www.youtube.com/watch?v=b9Uu7gRF5KY),
(2020),
[R Consortium at YouTube](https://www.youtube.com/channel/UC_R5smHVXRYGhZYDJsnXTwg).
