---
title: "QRMon data cleaning examples"
author: "Anton Antonov"
date: "1/22/2019"
output: html_notebook
---

```{r}
library(magrittr)
library(ggplot2)
library(quantreg)
library(splines)
library(QRMon)
```

# Introduction

In this notebook we consider couple of examples of cleaning data by 
(repeatedly) selecting data around regression quantiles.



# Deleting noisy signals from a plot manually and then export the best remaining signal

Data and cleaning approach taken from the Mathematica Stackexchange discussion:

["Deleting noisy signals from a plot manually and then export the best remaining signal"](https://mathematica.stackexchange.com/questions/188361/deleting-noisy-data-from-a-plot-manually-and-export-the-best-remaining-data).

```{r}
df <- read.csv("~/MathFiles/Questions/MSE-q188361.txt", header = FALSE, as.is = TRUE )
dim(df)
df <- df[complete.cases(df[,1:4]), 1:4]
df <- Reduce( function(a,i) { a[,i] <- as.numeric(a[,i]); a }, init = df, x = 1:ncol(df) )
dim(df)
summary(df)
```

```{r}
df2 <- 
  purrr::map_df( 1:ncol(df), 
              function(i) {
                data.frame( Regressor = 1:nrow(df), Value = df[,i])
              })
```

```{r}
df2
```

```{r, results='asis'}
obj <- QRMonUnit(df) %>% QRMonEchoDataSummary()
```


```{r}
obj2 <-
  QRMonUnit( df2) %>% 
  QRMonQuantileRegression(18, probabilities = 0.3) %>% 
  QRMonPlot
```

```{r, collapse=T}
res <-
  Reduce( 
    function(df, par) {
      cleanData <-
        QRMonUnit(df) %>% 
        QRMonEchoDataSummary() %>% 
        QRMonQuantileRegression( df = 30, probabilities = par$QuantileFraction, degree = 3 ) %>% 
        QRMonPlot(echoQ = T) %>% 
        QRMonErrors( relativeErrorsQ = FALSE ) %>% 
        QRMonEchoFunctionValue( function(x) map(x, summary) ) %>% 
        QRMonErrorsPlot( relativeErrorsQ = FALSE ) %>% 
        QRMonPickPathPoints( par$Threshold ) %>% 
        QRMonTakeValue()
      cleanData <- cleanData[[1]]
      print(ggplot(cleanData) + geom_point( aes( x = Regressor, y = Value ) ))
      cleanData
    },
    init = df2,
    x = list( list( QuantileFraction = 0.3, Threshold = 0.1 ), list( QuantileFraction = 0.5, Threshold = 0.025 ) ))
```

# Electrical measurements data

Here we repeat the procedure with another dataset from the MSE question 
["Remove periodic data variation from experimental data"](https://mathematica.stackexchange.com/q/194782/34008).

```{r}
data <- read.csv( "~/MathFiles/Questions/qMSE-194782-Iamorph.txt", sep = "\t", header = F)
summary(data)
```

```{r, collapse=T}
res <-
  Reduce( 
    function(df, par) {
      cleanData <-
        QRMonUnit(df) %>% 
        QRMonEchoDataSummary() %>% 
        QRMonQuantileRegression( df = par$Knots, probabilities = par$QuantileFraction, degree = 3 ) %>% 
        QRMonPlot(echoQ = T) %>% 
        QRMonErrors( relativeErrorsQ = FALSE ) %>% 
        QRMonEchoFunctionValue( function(x) map(x, summary) ) %>% 
        QRMonErrorsPlot( relativeErrorsQ = FALSE ) %>% 
        QRMonPickPathPoints( par$Threshold ) %>% 
        QRMonTakeValue()
      cleanData <- cleanData[[1]]
      print(ggplot(cleanData) + geom_point( aes( x = Regressor, y = Value ) ))
      cleanData
    },
    init = data,
    x = list( list( Knots = 30, QuantileFraction = 0.5, Threshold = 0.5 ), list( Knots = 50, QuantileFraction = 0.5, Threshold = 0.25 ) ))
```


```{r}
lattice::xyplot( Value ~ Regressor, res, type = c("g","p"), cex = 0.05 )
```

